Ainur:~A Framework for Repeatable End-to-End Wireless Edge Computing Testbed Research

### 2022-12-19 publications/2022Ainur/assumptions.tex
glossaries

\subsubsection{Time synchronization}
Hosts on the testbed synchronize their clocks to a central \ac{NTP} server.
This is however only required for consistent time measurements \emph{across} devices; measurements with respect to the clock of a single device (e.g.\ \ac{RTT} measurements) do not require synchronization.

@@

    The control network, whose links are represented in orange in \cref{fig:network}, carries control plane data between the control host and the rest of the hosts on the testbed.
    It also provides internet access through a gateway router, and control plane access to the cloud instances through the control \ac{VPN} gateway (more details on this below).

@@

    This network carries data generated by the workloads deployed on the testbed, and is represented by blue links in \cref{fig:network}
    It also provides access to the workload data \ac{VPN} link to cloud instances through the workload data gateway.
    Although the physical links between devices are always present, interfaces on this network by default have no \ac{IP} addresses --- these are managed dynamically by Ainur.

@@

    In order to manage it through software, the managed switch is also connected to the control network through a single link.
    The associated interface is on a separate, exclusive \ac{VLAN}, and no data from the workload data network is allowed to enter the control network.

    \item[\ac{VPN} gateways]

    Two \ac{VPN} gateways are present on the testbed; one for control traffic, and one for workload data traffic to \ac{AWS} cloud instances.
    More details on these gateways are given in \cref{sec:layer3}.

@@

\begin{description}[style=unboxed]
    % \item[\ac{APT}]

    % Hosts use \ac{APT} to manage installed packages.

@@

    Docker\footnote{\url{https://docs.docker.com/}} is installed on the hosts.
    It is configured to allow regular users to run the Docker client and the server is set up to listen on unencrypted \ac{TCP} port \num{2375} on all interfaces, in addition to the default local UNIX socket.
\end{description}

@@

    The \ac{SSH} client on the control host and the \ac{SSH} servers on the controlled hosts (i.e.\ workload hosts, \ac{VPN} gateways) have been configured to allow password-less, public-key based authentication.

@@

Cloud integration in Ainur is done in \ac{AWS}\footnote{\url{https://aws.amazon.com/}} and the associated \texttt{boto3}\footnote{\url{https://aws.amazon.com/sdk-for-python/}} Python library.

\begin{description}[style=unboxed]
    \item[Authentication to \ac{AWS}]

    Credentials to \ac{AWS} exist in the execution environment of Ainur on the control hosts.
    Specifically, these are credentials that allow Ainur to fully manage:
    \begin{itemize}
        \item \ac{EC2} instances, to create and terminate required compute nodes on-the-fly;
        \item security groups, to allow remote access over \ac{SSH} and \ac{VPN} to the compute nodes;
        \item and private access keys, to securely log in to the nodes over \ac{SSH}.
    \end{itemize}

@@

    \Ac{EC2} instances deployed by Ainur are expected to be based on a Linux \ac{AMI} equipped with the similar software as the local hosts.
    Specifically:

    \begin{itemize}
        \item Docker is installed and configured for regular users and over-the-network access on \ac{TCP} port \num{2375}.
        \item Netplan and systemd are present and in use.
        \item VPNCloud\footnote{\url{https://vpncloud.ddswd.de/}} is installed (see below).
    \end{itemize}

    \item[\ac{VPN} gateways]

@@

        Two gateways are present on the testbed networks, as illustrated in \cref{fig:network}; one for the control traffic \ac{VPN} link, and another for the workload data traffic \ac{VPN} link.

        \item[Public \ac{IP} address]

        Both gateways are reachable on the same public \ac{IP} address of the gateway router.
        The gateway router listens on two different \ac{UDP} ports and forwards data to the corresponding \ac{VPN} gateway.

@@

        Static routes to the control \ac{VPN} link through the corresponding gateway are configured on all devices attached to the control network.

>>>>

\subsubsection{Time synchronization}
Hosts on the testbed synchronize their clocks to a central \gls{NTP} server.
This is however only required for consistent time measurements \emph{across} devices; measurements with respect to the clock of a single device (e.g.\ \gls{RTT} measurements) do not require synchronization.

@@

    The control network, whose links are represented in orange in \cref{fig:network}, carries control plane data between the control host and the rest of the hosts on the testbed.
    It also provides internet access through a gateway router, and control plane access to the cloud instances through the control \gls{VPN} gateway (more details on this below).

@@

    This network carries data generated by the workloads deployed on the testbed, and is represented by blue links in \cref{fig:network}
    It also provides access to the workload data \gls{VPN} link to cloud instances through the workload data gateway.
    Although the physical links between devices are always present, interfaces on this network by default have no \gls{IP} addresses --- these are managed dynamically by Ainur.

@@

    In order to manage it through software, the managed switch is also connected to the control network through a single link.
    The associated interface is on a separate, exclusive \gls{VLAN}, and no data from the workload data network is allowed to enter the control network.

    \item[\gls{VPN} gateways]

    Two \gls{VPN} gateways are present on the testbed; one for control traffic, and one for workload data traffic to \gls{AWS} cloud instances.
    More details on these gateways are given in \cref{sec:layer3}.

@@

\begin{description}[style=unboxed]
    % \item[\gls{APT}]

    % Hosts use \gls{APT} to manage installed packages.

@@

    Docker\footnote{\url{https://docs.docker.com/}} is installed on the hosts.
    It is configured to allow regular users to run the Docker client and the server is set up to listen on unencrypted \gls{TCP} port \num{2375} on all interfaces, in addition to the default local UNIX socket.
\end{description}

@@

    The \gls{SSH} client on the control host and the \gls{SSH} servers on the controlled hosts (i.e.\ workload hosts, \gls{VPN} gateways) have been configured to allow password-less, public-key based authentication.

@@

Cloud integration in Ainur is done in \gls{AWS}\footnote{\url{https://aws.amazon.com/}} and the associated \texttt{boto3}\footnote{\url{https://aws.amazon.com/sdk-for-python/}} Python library.

\begin{description}[style=unboxed]
    \item[Authentication to \gls{AWS}]

    Credentials to \gls{AWS} exist in the execution environment of Ainur on the control hosts.
    Specifically, these are credentials that allow Ainur to fully manage:
    \begin{itemize}
        \item \gls{EC2} instances, to create and terminate required compute nodes on-the-fly;
        \item security groups, to allow remote access over \gls{SSH} and \gls{VPN} to the compute nodes;
        \item and private access keys, to securely log in to the nodes over \gls{SSH}.
    \end{itemize}

@@

    \gls{EC2} instances deployed by Ainur are expected to be based on a Linux \gls{AMI} equipped with the similar software as the local hosts.
    Specifically:

    \begin{itemize}
        \item Docker is installed and configured for regular users and over-the-network access on \gls{TCP} port \num{2375}.
        \item Netplan and systemd are present and in use.
        \item VPNCloud\footnote{\url{https://vpncloud.ddswd.de/}} is installed (see below).
    \end{itemize}

    \item[\gls{VPN} gateways]

@@

        Two gateways are present on the testbed networks, as illustrated in \cref{fig:network}; one for the control traffic \gls{VPN} link, and another for the workload data traffic \gls{VPN} link.

        \item[Public \gls{IP} address]

        Both gateways are reachable on the same public \gls{IP} address of the gateway router.
        The gateway router listens on two different \gls{UDP} ports and forwards data to the corresponding \gls{VPN} gateway.

@@

        Static routes to the control \gls{VPN} link through the corresponding gateway are configured on all devices attached to the control network.

### 2022-12-19 publications/2022Ainur/conclusion.tex
glossaries

In this paper, we have introduced a framework for repeatable end-to-end testbed automation in the context of wireless networking and edge computing research.
Named Ainur, it simplifies the execution and verification of end-to-end experimentation by automating the
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item establishment of physical links between hosts, including the configuration of complex wireless systems such as 4G \ac{LTE} and 5G
    \item provisioning of and connection to remote cloud instances
    \item initialization of \ac{IP} layer connectivity between hosts
    \item collection of logs and data
    \item deployment, scaling, and lifecycle management of containerized processes
\end{enumerate*}.
We have described its general architecture, which follows a layered design mimicking the network stack layers the framework directly interacts with, as well as the underlying assumptions about its deployment environment and specific requirements for its deployment.
We have also outlined a demonstration which showcases the flexibility and power of the framework by deploying two different workloads to our testbed.
We believe our framework represents an important step towards repeatable, replicable, yet low-access barrier end-to-end wireless testbed experimentation.
It has been released as \ac{FOSS} and can be found on GitHub~\cite{ainur:github}.\\

In terms of future work, our current efforts are focused on the expansion and integration of Ainur with \ac{CHI}~\cite{keahey2020lessons}.
While we believe Ainur to be unique and valuable in its focus on fully-automated end-to-end wireless edge computing experimentation, integrating with the above software stacks will provide a number of features and functionalities which will greatly expand Ainur's potential.
Features like resource reservation, multi-tenant testbed access, and automated bare-metal and \ac{VM}-host instantiation will make Ainur better suited for larger scale experimentation and allow us to expand the research domain targeted by the framework.
Our ultimate goal is to eventually provide Ainur as a service running inside an OpenStack/\ac{CHI} environment, leveraging the flexibility of the platform to automate the reservation of resources, instantiation of nodes and networks, execution of experiments, and final collection of results.

>>>>

In this paper, we have introduced a framework for repeatable end-to-end testbed automation in the context of wireless networking and edge computing research.
Named Ainur, it simplifies the execution and verification of end-to-end experimentation by automating the
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item establishment of physical links between hosts, including the configuration of complex wireless systems such as 4G \gls{LTE} and 5G
    \item provisioning of and connection to remote cloud instances
    \item initialization of \gls{IP} layer connectivity between hosts
    \item collection of logs and data
    \item deployment, scaling, and lifecycle management of containerized processes
\end{enumerate*}.
We have described its general architecture, which follows a layered design mimicking the network stack layers the framework directly interacts with, as well as the underlying assumptions about its deployment environment and specific requirements for its deployment.
We have also outlined a demonstration which showcases the flexibility and power of the framework by deploying two different workloads to our testbed.
We believe our framework represents an important step towards repeatable, replicable, yet low-access barrier end-to-end wireless testbed experimentation.
It has been released as \gls{FOSS} and can be found on GitHub~\cite{ainur:github}.\\

In terms of future work, our current efforts are focused on the expansion and integration of Ainur with \gls{CHI}~\cite{keahey2020lessons}.
While we believe Ainur to be unique and valuable in its focus on fully-automated end-to-end wireless edge computing experimentation, integrating with the above software stacks will provide a number of features and functionalities which will greatly expand Ainur's potential.
Features like resource reservation, multi-tenant testbed access, and automated bare-metal and \gls{VM}-host instantiation will make Ainur better suited for larger scale experimentation and allow us to expand the research domain targeted by the framework.
Our ultimate goal is to eventually provide Ainur as a service running inside an OpenStack/\gls{CHI} environment, leveraging the flexibility of the platform to automate the reservation of resources, instantiation of nodes and networks, execution of experiments, and final collection of results.

### 2022-12-19 publications/2022Ainur/demo.tex
glossaries

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022Ainur/figures/demo_hardware.png}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
            \item \ac{SDR} processing hosts 
            \item additional base-station host for the Core Network and \ac{eNodeB}
        \end{enumerate*}.
    }\label{fig:democonfigs}
\end{figure*}

@@

\Cref{fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \ac{AWS} \ac{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \iac{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board WiFi.
    \item[4G \ac{LTE}:] \iac{SDR} is configured as an 4G \ac{LTE} base station, and another is configured as an 4G \ac{LTE} \ac{UE}.
    Together these radios act as an \ac{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \ac{AWS}'s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{fig:wkld}.
The first of these (\cref{fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \ac{MNIST} dataset~\cite{mnist}.
The application consists of a client with a web interface, and \iac{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{fig:wkld:ncs}) corresponds to \iac{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \acp{NCS} using cloud-native technologies~\cite{cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \ac{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \ac{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \ac{VPN} gateway for the management network, as well as hosting the necessary \ac{NTP} and \ac{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \ac{AWS} \ac{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \acp{SDR}, and \ac{VPN} gateways; please refer to \cref{fig:network} for an architectural overview of this setup.

@@

Participants will be asked to choose
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \ac{AWS} datacenter)
    \item a physical layer for the first hop of the network (WiFi, 4G \ac{LTE}, or ethernet). 
\end{enumerate*}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \ac{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{fig:mockup}.

@@

Participants will be asked to specify the full, arbitrarily complex, experimental scenario, including the number of clients, physical layers (single or multiple, and which clients are on each physical link), and offloading configuration (number of instances on the edge vs.\ on the cloud, which \ac{AWS} datacenters to deploy to).
Configuration will be specified through a \acs{YAML} file, which will then be parsed by Ainur for the automatic execution of the scenario.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022Ainur/figures/demo_hardware.png}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
            \item \gls{SDR} processing hosts
            \item additional base-station host for the Core Network and \gls{eNodeB}
        \end{enumerate*}.
    }\label{fig:democonfigs}
\end{figure*}

@@

\Cref{fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \iac{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board WiFi.
    \item[4G \gls{LTE}:] \iac{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \gls{AWS}'s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{fig:wkld}.
The first of these (\cref{fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{mnist}.
The application consists of a client with a web interface, and \iac{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{fig:wkld:ncs}) corresponds to \iac{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \acp{NCS} using cloud-native technologies~\cite{cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \acp{SDR}, and \gls{VPN} gateways; please refer to \cref{fig:network} for an architectural overview of this setup.

@@

Participants will be asked to choose
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \gls{AWS} datacenter)
    \item a physical layer for the first hop of the network (WiFi, 4G \gls{LTE}, or ethernet).
\end{enumerate*}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \gls{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{fig:mockup}.

@@

Participants will be asked to specify the full, arbitrarily complex, experimental scenario, including the number of clients, physical layers (single or multiple, and which clients are on each physical link), and offloading configuration (number of instances on the edge vs.\ on the cloud, which \gls{AWS} datacenters to deploy to).
Configuration will be specified through a \acs{YAML} file, which will then be parsed by Ainur for the automatic execution of the scenario.

### 2022-12-19 publications/2022Ainur/futurework.tex
glossaries

    \begin{enumerate}
        \item Components in the implementation are quite tightly coupled, so adding new functionality or layers in many cases implies rewritting large chunks of code.
        This also requires quite a lot of synchronization between people working on different parts of the code.
        \item It makes experimentation synchronous, and when running things over \ac{SSH}, it requires the use of tools such as \texttt{screen} or \texttt{tmux} to be able to run long experiments without having to keep the connection open at all times.
        \item We currently have no way of ensuring mutual exclusion of multiple instances of Ainur on the testbed, and implementing such functionality would probably require some sort of centralized instance tracking mechanism.
    \end{enumerate}

    I (Manuel) suggest then that for ExPECA 2.0 we switch to a distributed client-server architecture, using a microservice for each core functionality.
    In other words, we would have several smaller server applications running on potentially different hosts on the network, each handling a specific part of the functionality we need.
    For instance, we could have a ``Layer 3'' service which handles setting up \ac{IP}-layer connectivity, and another ``SDR'' service which manages the \acp{SDR}.
    Services could communicate with each other using simple HTTP \acp{API} for easy debugging, and we would have an ``Ainur'' client which orchestrates the services into an experimental run.

@@

    \begin{enumerate}
        \item It decouples the implementation details for each service.
        As long as the \ac{API} is well-defined, it doesn't matter how the service is implemented - we could even use different languages for different functionalities.
        \item Adding new functionality becomes simply a matter of writing a new microservice and plugging it into the existing setup.
        \item Experimentation can be done asynchronously through an ``experiment scheduler'' service which queues, runs, and then collects results from experiments.
        This service would also then ensure mutual exclusion of testbed resources.
        \item Things could be parallelized.
        Our current design is sequential, even when it doesn't need to be.
        Things that do not depend on each other could be executed in parallel (e.g. cloud and local configuration), cutting down on experiment time overhead.
        \item It would fit squarely into the design of both Docker and Kubernetes, both of which are distributed client-service architectures.
    \end{enumerate}

>>>>

    \begin{enumerate}
        \item Components in the implementation are quite tightly coupled, so adding new functionality or layers in many cases implies rewritting large chunks of code.
        This also requires quite a lot of synchronization between people working on different parts of the code.
        \item It makes experimentation synchronous, and when running things over \gls{SSH}, it requires the use of tools such as \texttt{screen} or \texttt{tmux} to be able to run long experiments without having to keep the connection open at all times.
        \item We currently have no way of ensuring mutual exclusion of multiple instances of Ainur on the testbed, and implementing such functionality would probably require some sort of centralized instance tracking mechanism.
    \end{enumerate}

    I (Manuel) suggest then that for ExPECA 2.0 we switch to a distributed client-server architecture, using a microservice for each core functionality.
    In other words, we would have several smaller server applications running on potentially different hosts on the network, each handling a specific part of the functionality we need.
    For instance, we could have a ``Layer 3'' service which handles setting up \gls{IP}-layer connectivity, and another ``SDR'' service which manages the \acp{SDR}.
    Services could communicate with each other using simple HTTP \acp{API} for easy debugging, and we would have an ``Ainur'' client which orchestrates the services into an experimental run.

@@

    \begin{enumerate}
        \item It decouples the implementation details for each service.
        As long as the \gls{API} is well-defined, it doesn't matter how the service is implemented - we could even use different languages for different functionalities.
        \item Adding new functionality becomes simply a matter of writing a new microservice and plugging it into the existing setup.
        \item Experimentation can be done asynchronously through an ``experiment scheduler'' service which queues, runs, and then collects results from experiments.
        This service would also then ensure mutual exclusion of testbed resources.
        \item Things could be parallelized.
        Our current design is sequential, even when it doesn't need to be.
        Things that do not depend on each other could be executed in parallel (e.g. cloud and local configuration), cutting down on experiment time overhead.
        \item It would fit squarely into the design of both Docker and Kubernetes, both of which are distributed client-service architectures.
    \end{enumerate}

### 2022-12-19 publications/2022Ainur/introduction.tex
glossaries

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \acs*{COSMOS}
    \item \acs*{POWDER}
    \item \acs*{ARA}
    \item Drexel Grid \ac{SDR}
\end{enumerate*} testbeds.
\acs{COSMOS}~(\emph{\acl{COSMOS}})\acused{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \acp{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{Cosmos1,cosmos2}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\ac{COSMOS} relies on the \ac{OMF} (originally developed for ORBIT~\cite{orbit}), and employs the \ac{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\acs{POWDER}~(\emph{\acl{POWDER}})\acused{POWDER} promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \acp{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes. 
\ac{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \ac{VM} images containing the necessary software and configurations.
These profiles are defined through using \ac{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

The \ac{ARA} platform is an at-scale testbed for wireless research spanning a rural area with a diameter of over \SI{60}{\kilo\meter} in Iowa~\cite{zhang2022ara}.
Its core goal is the study and deployment of advanced wireless platforms and technologies in real-world agricultural and rural settings.
It includes a broad range of wireless platforms ranging from low-\ac{UHF} massive \ac{MIMO} to \unit{\milli\meter}Wave, deployed through both \acp{SDR} and programmable \ac{COTS} radios, as well as automated ground vehicles, cameras and sensors.
\ac{ARA}'s software stack, ARASoft, is based upon the highly flexible and powerful \ac{CHI} software suite from the Chameleon testbed project~\cite{keahey2020lessons}, which in turn is based on the widely adopted OpenStack~\cite{openstack} cloud-computing framework.
This affords the \ac{ARA} testbed a high degree of flexibility, as well as lowers the learning curve for new contributors and users.

Finally, the \emph{Drexel Grid \ac{SDR} Testbed} features \acp{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{DrexelGrid}.
Primarily intended for \ac{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \ac{LXC} runtime for the deployment of both experimental code and \ac{SDR} software, which affords users great freedom when it comes to the development of experiments.

Experimentation is key to to fully understanding the implications of next-generation wireless systems, cloud, and edge computing paradigms, and thus more of these testbeds are sure to emerge in the near future.
Yet, little work has so far been devoted to general-purpose, hardware-agnostic software frameworks for the management and automation of such systems.
Existing platforms implement their own, ad-hoc software solutions which are not compatible with other testbeds, and in many cases are not even compatible with reigning cloud-native standards.
This is, for instance, the case with \ac{COSMOS} and \ac{POWDER}; their reliance on domain-specific languages limits their integration with cloud-native solutions, which generally build upon general-purpose languages such as Python and Go.
These testbeds further leverage virtualization technology based on \acp{VM} instead of more lightweight and edge-compatible solutions such as containers.

@@

In this work, we present our solution to the challenge of testbed automation: Ainur, a framework for wireless testbed automation with a specific focus on end-to-end experimental research in the context of edge-computing using cloud- and edge-native technologies.
Ainur is designed to deploy experimental runs from a workload perspective by configuring the physical testbed, initializing all involved software components, deploying and executing the experimental workload, collecting logs and data, and finally gracefully degrading the system.
The framework allows for dynamic, software-definition of physical and logical links, network topology, cloud and edge computing resources, as well as experimental workload deployment and orchestration.
It heavily leverages cloud-native technologies, such as Docker containers, in order to support a wide variety of different testbed hardware setups and experimental configurations and workloads, as well as to be as easily extendable as possible.
Furthermore, we make Ainur available to the community as \ac{FOSS}.
It can be obtained from the {KTH-EXPECA/Ainur} repository on GitHub~\cite{ainur:github}, released under an Apache version \num{2.0} license.

@@

% \todo[inline]{
%     COSMOS: Experimenters work on these resources using the \ac*{OMF}~\cite{omf} tool, originally leveraged for ORBIT~\cite{orbit}, and \ac{SSH} sessions, thus allowing for free-form scripting of experimental procedures.

@@

%     POWDER: To abstract and redistribute resources and \acp{API}, \ac{POWDER} leverages the concept of \emph{profiles}, corresponding shareable packages containing experimental configurations and descriptions\cite{emulab}. 
%     Automation and web interface control are limited to some test experiments, and thus regular users must interact with the platform through \ac{SSH}.

@@

%     Drexel: Orchestration in this testbed is facilitated through a container-based approach, leveraging LXC containers, while deployment and configuration is facilitated through Ansible.
%     Container deployment is requested by users through a web-based reservation system or a \ac{REST} web \ac{API}.
%     \ac{OS} deployment on host computers is facilitated through Canonical \ac{MaaS}.

>>>>

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \acs*{COSMOS}
    \item \acs*{POWDER}
    \item \acs*{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\acs{COSMOS}~(\emph{\acl{COSMOS}})\acused{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \acp{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{Cosmos1,cosmos2}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

\acs{POWDER}~(\emph{\acl{POWDER}})\acused{POWDER} promises research on wireless and mobile networks with a level of programmability down to the waveform~\cite{powder}.
The testbed spans a \SI{15}{\kilo\meter\squared} area and features about \num{15} fixed programmable radio nodes, based on off-the-shelves \acp{SDR} and featuring edge-like compute capabilities and integration with cloud resources, which interact with \num{50} mobile nodes. 
\gls{POWDER} experiments are defined and developed in \emph{profiles}, which correspond to \gls{VM} images containing the necessary software and configurations.
These profiles are defined through using \gls{RSpec}\footnote{\url{https://groups.geni.net/geni/wiki/GENIExperimenter/RSpecs}} documents.

The \gls{ARA} platform is an at-scale testbed for wireless research spanning a rural area with a diameter of over \SI{60}{\kilo\meter} in Iowa~\cite{zhang2022ara}.
Its core goal is the study and deployment of advanced wireless platforms and technologies in real-world agricultural and rural settings.
It includes a broad range of wireless platforms ranging from low-\gls{UHF} massive \gls{MIMO} to \unit{\milli\meter}Wave, deployed through both \acp{SDR} and programmable \gls{COTS} radios, as well as automated ground vehicles, cameras and sensors.
\gls{ARA}'s software stack, ARASoft, is based upon the highly flexible and powerful \gls{CHI} software suite from the Chameleon testbed project~\cite{keahey2020lessons}, which in turn is based on the widely adopted OpenStack~\cite{openstack} cloud-computing framework.
This affords the \gls{ARA} testbed a high degree of flexibility, as well as lowers the learning curve for new contributors and users.

Finally, the \emph{Drexel Grid \gls{SDR} Testbed} features \acp{SDR} that connect either over-the-air, through a channel emulator, or over a combination of the two, to facilitate realistic and reproducible experimentation~\cite{DrexelGrid}.
Primarily intended for \gls{SDR}-centric research, it does not integrate any core, cloud or edge components.
However, the testbed extensively employs the \gls{LXC} runtime for the deployment of both experimental code and \gls{SDR} software, which affords users great freedom when it comes to the development of experiments.

Experimentation is key to to fully understanding the implications of next-generation wireless systems, cloud, and edge computing paradigms, and thus more of these testbeds are sure to emerge in the near future.
Yet, little work has so far been devoted to general-purpose, hardware-agnostic software frameworks for the management and automation of such systems.
Existing platforms implement their own, ad-hoc software solutions which are not compatible with other testbeds, and in many cases are not even compatible with reigning cloud-native standards.
This is, for instance, the case with \gls{COSMOS} and \gls{POWDER}; their reliance on domain-specific languages limits their integration with cloud-native solutions, which generally build upon general-purpose languages such as Python and Go.
These testbeds further leverage virtualization technology based on \acp{VM} instead of more lightweight and edge-compatible solutions such as containers.

@@

In this work, we present our solution to the challenge of testbed automation: Ainur, a framework for wireless testbed automation with a specific focus on end-to-end experimental research in the context of edge-computing using cloud- and edge-native technologies.
Ainur is designed to deploy experimental runs from a workload perspective by configuring the physical testbed, initializing all involved software components, deploying and executing the experimental workload, collecting logs and data, and finally gracefully degrading the system.
The framework allows for dynamic, software-definition of physical and logical links, network topology, cloud and edge computing resources, as well as experimental workload deployment and orchestration.
It heavily leverages cloud-native technologies, such as Docker containers, in order to support a wide variety of different testbed hardware setups and experimental configurations and workloads, as well as to be as easily extendable as possible.
Furthermore, we make Ainur available to the community as \gls{FOSS}.
It can be obtained from the {KTH-EXPECA/Ainur} repository on GitHub~\cite{ainur:github}, released under an Apache version \num{2.0} license.

@@

% \todo[inline]{
%     COSMOS: Experimenters work on these resources using the \ac*{OMF}~\cite{omf} tool, originally leveraged for ORBIT~\cite{orbit}, and \gls{SSH} sessions, thus allowing for free-form scripting of experimental procedures.

@@

%     POWDER: To abstract and redistribute resources and \acp{API}, \gls{POWDER} leverages the concept of \emph{profiles}, corresponding shareable packages containing experimental configurations and descriptions\cite{emulab}.
%     Automation and web interface control are limited to some test experiments, and thus regular users must interact with the platform through \gls{SSH}.

@@

%     Drexel: Orchestration in this testbed is facilitated through a container-based approach, leveraging LXC containers, while deployment and configuration is facilitated through Ansible.
%     Container deployment is requested by users through a web-based reservation system or a \gls{REST} web \gls{API}.
%     \gls{OS} deployment on host computers is facilitated through Canonical \gls{MaaS}.
